cardinalR: Generating interesting high-dimensional data structures

### 2025-04-25 paper-cardinalR.Rmd
Add intro to the first application

```{r load-libraries}
library(cardinalR)
library(tidyverse)
library(kableExtra)
```

@@

## Assessing the performance of dimension reduction on different geomatric structures in high-dimensions

We illustrate the use of the package by generating five clusters of different shapes, positioned with equal inter-cluster distances in $4\text{-}D$. Using this synthetic dataset, we will evaluate the performance of three popular dimensionality reduction techniques: Principal Component Analysis (PCA), t-distributed Stochastic Neighbor Embedding (tSNE), and Uniform Manifold Approximation and Projection (UMAP).

>>>>

```{r load-libraries}
library(cardinalR)
library(tidyverse)
library(kableExtra)
library(geozoo)
```

@@

## Assessing the performance of dimension reduction on different geometric structures in high-dimensions

We illustrate the use of the package by generating five clusters of different shapes, positioned with equal inter-cluster distances in $4\text{-}D$. Using this synthetic dataset, we will evaluate the performance of six popular dimensionality reduction techniques: Principal Component Analysis (PCA) [@jolliffe2011], t-distributed stochastic neighbor embedding (tSNE) [@laurens2008], uniform manifold approximation and projection (UMAP) [@leland2018], potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm [@moon2019], large-scale dimensionality reduction Using triplets (TriMAP) [@amid2019], and pairwise controlled manifold approximation (PaCMAP) [@yingfan2021].
